---
title: "Surrogates"
author: "Jinshi"
date: "July 15, 2019"
output: html_document
---

```{r message=FALSE}
# install.packages('kableExtra')
# install.packages('ggpubr')
library(data.table)
library(lubridate)
library(kableExtra)
library(gridExtra)
library(cowplot)
library(knitr)
library("ggpubr")
# install.packages('tidyverse')
library(tidyverse)
# library(reshape)
# install.packages("ggmap")
library(ggmap)
# install.packages("maps")
library(maps)
# install.packages("mapdata")
library(mapdata)
# library(tidyr)
# install.packages("hexbin")
library(hexbin)
library(cowplot)
# Source all needed functions
library(ggplot2)
theme_set(theme_bw())
# bi-plot
# install.packages("devtools")
library(devtools)
# install_github("vqv/ggbiplot")
# library(ggbiplot)
library(rpart)
source('functions.R')
library(dplyr)
# install.packages('egg', dependencies = TRUE)
library(egg)

```




```{r preliminaries, message=FALSE, include=FALSE, echo=FALSE, cache=TRUE}
# Set chunks defaults; these options will be applied to all subsequent chunks
knitr::opts_chunk$set(results = 'hide', message = TRUE, include = TRUE, echo = FALSE,
                      fig.height = 4, fig.width = 8, cache = TRUE)
# Constants
OUTPUT_DIR		<- "outputs/"
DATA_DIR <- 'data'
# Create output and log folders if they do not exist
if(!file.exists(OUTPUT_DIR)) dir.create(OUTPUT_DIR)
```


```{r}
SRDB <- read.csv(file.path(DATA_DIR,"srdb-data.csv"))
DGRsD <- read.csv(file.path(DATA_DIR, "DGRsD.csv"))
DGRsD %>% filter(is.na(SiteID))

DGRsD %>% filter(Climate %in% c("Af", "Am", "As", "Aw"))
DGRsD %>% mutate(Climate_type = case_when(
  Climate %in% c("Af", "Am", "As", "Aw") ~ "(A) Tropic",
  Climate %in% c("BSh", "BSk", "BWh", "BWk") ~ "(B) Arid",
  Climate %in% c("Dfa", "Dfb", "Dfc", "Dsc", "Dwa", "Dwb", "Dwc", "EF", "ET") ~ "(D) Continental", 
  TRUE ~ "(C) Temperate"
)) -> DGRsD

DGRsD %>% select (Climate, Climate_type) %>% unique() %>% arrange(Climate_type)

DGRsD %>% filter(!is.na(Tsoil) & !is.na(Tm_Del) & !is.na(RS_Norm)) -> DGRsD_TS
DGRsD %>% select(Climate) %>% unique()


DGRsD %>% filter(SWC_Type == "VWC" | SWC_Type == "GWC") %>% 
  filter (!is.na(SWC) & !is.na(Pm_Del) & !is.na(RS_Norm)) -> DGRsD_SWC

DGRsD_SWC$Ecosystem_type <- ifelse(is.na(DGRsD_SWC$Ecosystem_type), "Not available", DGRsD_TS$Ecosystem_type)
GPT <- read.csv(file.path(DATA_DIR, "summarized_climate.csv"))
```


```{r}
# SRDB clear Biome, Ecosystem, Leaf habitat, and drainage
SRDB %>% mutate(Biome_group = case_when(Biome == "Boreal" ~ "Boreal"
                                  ,Biome == "Temperate" ~ "Temperate"
                                  ,Biome == "Arctic" | Biome == "Polar" ~ "Arctic"
                                  ,Biome == "Tropical" ~ "Tropical"
                                  ,Biome == "Mediterranean" ~ "Mediterranean"
                                  ,Biome == "Subtropical" | Biome == "Subtropic" ~ "Subtropical"
                                  ,TRUE ~ "Not reported"
                                  )) -> SRDB
SRDB %>% select(Biome_group) %>% count(Biome_group)

SRDB %>% mutate(Ecosystem_group = case_when(Ecosystem_type == "" ~ "Not reported"
                                            , Ecosystem_type == "Bare" | Ecosystem_type == "Bare land" | Ecosystem_type == "Bare soil" ~ "BSV"
                                            , Ecosystem_type == "Desert" ~ "Desert"
                                            , Ecosystem_type %in% c("Forest", "Natural")  ~ "Forest"
                                            , Ecosystem_type %in% c("Garden", "Orchard", "Plantation")  ~ "Plantation"
                                            , Ecosystem_type %in% c("Grassland", "Pasture", "Meadow") ~ "Grassland"
                                            , Ecosystem_type == "Peatland" ~ "Peatland"
                                            , Ecosystem_type == "Savanna" ~ "Savanna"
                                            , Ecosystem_type %in% c("Shrubland", "Woodland")  ~ "Shrubland"
                                            , Ecosystem_type == "Tundra" ~ "Tundra"
                                            , Ecosystem_type == "Urban" ~ "Urban"
                                            , Ecosystem_type == "Wetland" ~ "Wetland"
                                            , TRUE ~ "Agriculture"
                                            )) -> SRDB
# unique(SRDB$Ecosystem_type) %>% sort

SRDB %>% select(Ecosystem_group, Ecosystem_type) %>% unique() %>% arrange(Ecosystem_group)


SRDB %>% mutate(Leaf_group = case_when (Leaf_habit == "Deciduous" ~ "Deciduous"
                                        , Leaf_habit %in% c("Evergreen", "Evergreen broadleaf")  ~ "Evergreen"
                                        , Leaf_habit %in% c("Mixed", "Mixed of evergreen and deciduous") ~ "Mixed"
                                        , TRUE ~ "Not reported"
                                        )) -> SRDB

# unique(SRDB$Leaf_habit) %>% sort
SRDB %>% select(Leaf_group, Leaf_habit) %>% unique() %>% arrange(Leaf_group)


SRDB %>% mutate(Drainage_group = case_when (Soil_drainage == "Dry" ~ "Dry"
                                        , Soil_drainage %in% c("Wet")  ~ "Wet"
                                        , Soil_drainage %in% c("Mixed", "Medium") ~ "Medium"
                                        , TRUE ~ "Not reported"
                                        )) -> SRDB
SRDB %>% select(Drainage_group, Soil_drainage) %>% unique() %>% arrange(Drainage_group)
# unique(SRDB$Soil_drainage) %>% sort
```



```{r}
# plot Figure1 - sites with Tsoil and SWC vs all data from DGRsD

ggplot(aes(Tm_Del, Pm_Del), data = DGRsD) + geom_bin2d(alpha = 0.5) + 
  scale_fill_gradient(low = "lightgray", high = "black") +
  xlab(expression(Temperature~"( "~degree~C~")")) +
  ylab(expression(Precipitation~"(mm"~month^{-1}~")")) +
  geom_jitter(data = DGRsD_TS, aes(Tm_Del, Pm_Del, col = Climate_type), alpha = 1) +
  theme (legend.position = c(0.25, 0.65), legend.background = element_rect(fill = alpha("white", 0), size = 0.75) ) +
  guides (fill = F) -> p1
  
ggplot(aes(Tm_Del, Pm_Del), data = DGRsD) + geom_bin2d(alpha = 0.5) + 
  scale_fill_gradient(low = "lightgray", high = "black") +
  xlab(expression(Temperature~"( "~degree~C~")")) +
  ylab(expression(Precipitation~"(mm"~month^{-1}~")")) +
  geom_jitter(data = DGRsD_SWC, aes(Tm_Del, Pm_Del, col = Climate_type), alpha = 0.15) +
  theme (legend.position = "none" ) +
  guides (fill = F) -> p2

plot_grid(p1, p2, labels = c( "(a)", "(b)" )
          , vjust = 2, hjust = c(-4), ncol = 2 )

ggsave("outputs/Figure1-1. Site distribution.jpg", width = 8, height = 4, dpi = 300, units = "in")

```



```{r}
# soil temperature depth 
ggplot(DGRsD) +
  aes(TsDepth) +
  geom_histogram(fill = "gray", col = "black", bins = 30) +
  # scale_fill_gradient(low = "gray", high = "blue") +
  # geom_smooth(method = "lm") + 
  xlab(expression(Tsoil~measure~depth~"("~cm~")")) +
  ylab(expression(count)) -> p1

# SWC depth 
ggplot(subset(DGRsD, DGRsD$SWC_Type == "VWC" | DGRsD$SWC_Type == "GWC")) +
  aes(SWC_Depth) +
  geom_histogram(fill = "gray", col = "black", bins = 30) +
  # scale_fill_gradient(low = "gray", high = "blue") +
  # geom_smooth(method = "lm") + 
  xlab(expression(SWC~measure~depth~"("~cm~")")) +
  ylab(expression(count)) -> p2


# Group R2_Tsoil - R2_Tair into ()
t_results %>% mutate(Diff = first_R2 - first_Tm_R2) -> t_results
t_results %>% ggplot(aes(x = paste0("Tsoil (n=920)"), y = Diff)) + geom_violin() +  
  geom_jitter(shape = 16, position = position_jitter(0.2), col = 'gray') +
  geom_boxplot(width = 0.1) + 
  ylab(expression(SLR~R^2~difference)) + 
  theme(axis.title.x = element_blank()
        , axis.text=element_text(size=12)
        , axis.title.y = element_text(size = 12)) -> p3

t_results %>% ggplot(aes(x = paste0("Tsoil (n=920)"), y = poly_R2 - poly_Tm_R2)) + geom_violin() +  
  geom_jitter(shape = 16, position = position_jitter(0.2), col = 'gray') +
  geom_boxplot(width = 0.1) + 
  ylab(expression(Poly~R^2~difference)) + 
  theme(axis.title.x = element_blank()
        , axis.text=element_text(size=12)
        , axis.title.y = element_text(size = 12)) -> p4


results %>% ggplot(aes(x = paste0("SWC (n=528)"), y = first_R2 - first_pm_R2)) + geom_violin() +  
  geom_jitter(shape = 16, position = position_jitter(0.2), col = 'gray') +
  geom_boxplot(width = 0.1) + 
  ylab(expression(SLR~R^2~difference)) + 
  theme(axis.title.x = element_blank()
        , axis.text=element_text(size=12)
        , axis.title.y = element_text(size = 12)) -> p5

results %>% ggplot(aes(x = paste0("SWC (n=528)"), y = poly_R2 - poly_pm_R2)) + geom_violin() +  
  geom_jitter(shape = 16, position = position_jitter(0.2), col = 'gray') +
  geom_boxplot(width = 0.1) + 
  ylab(expression(Poly~R^2~difference)) + 
  theme(axis.title.x = element_blank()
        , axis.text=element_text(size=12)
        , axis.title.y = element_text(size = 12)) -> p6


quantile(t_results$Diff)
quantile(t_results$poly_R2 - t_results$poly_Tm_R2)
# t_results %>% transmute(poly_diff = poly_R2 - poly_Tm_R2) 

quantile(results$first_R2 - results$first_pm_R2, na.rm = T)
quantile(results$poly_R2 - results$poly_pm_R2, na.rm = T)

plot_grid (p1, p3, p4, p2, p5, p6, ncol = 3
           , labels = c("(a)", "(b)", "(c)", "(d)", "(e)", "(f)")
           , hjust = c(-3.75, -3.75,-3.85, -3.75, -3.75, -5), vjust = c(2))

ggsave("outputs/Figure S1.jpg", dpi = 300, width = 8, height = 6, units = "in")

```



  
  
```{r, results='hide', message=FALSE, include=FALSE}
# model statistics
# t_results_orig <- t_model_sum()
t_results <- t_results_orig

t_results %>% na.omit() -> t_results
# t_results %>% rename (Study_number = sub_site.Study_number.1., SiteID = var_site.j.) -> t_results
colnames(t_results)[1:3] <- c("Study_number", "Measure_year", "SiteID")
# t_results %>% select (var_site.j.,  sub_site.Study_number.1.) 
t_results %>% select(Study_number, Measure_year, SiteID) %>% unique()
```




```{r}
# group R2 difference into three groups: low (Tsoil explianed less RS variability), mid (same), and high (Tsoil explained more RS variability)
t_results %>% mutate(Diff = first_R2 - first_Tm_R2) -> t_results
t_results <- set_r2_dif_group(low = -0.1, high = 0.1, sdata = t_results)

# use left join and get Ts_depth information
DGRsD_TS %>% 
  select(Study_number, SiteID, TsDepth, Latitude, Longitude, MAP_Del, MAT_Del) %>% group_by(Study_number, SiteID) %>% 
  dplyr::summarise(Ts_depth = mean(TsDepth), Lat = mean(Latitude), Lon = mean(Longitude), MAP = mean(MAP_Del), MAT = mean(MAT_Del)) %>%
  right_join(t_results, by = c("Study_number", "SiteID")) -> t_results
 
# more numeric factors from SRDB_V4
SRDB %>% select(Study_number, SiteID, Elevation) %>% group_by (Study_number, SiteID) %>%
 dplyr:: summarise(Ele = mean(Elevation, na.rm = TRUE)) %>% 
  right_join(t_results, by = c("Study_number", "SiteID")) -> t_results

# more caetgorial factors from SRDB_V4
SRDB %>% select(Study_number, SiteID, Biome_group, Ecosystem_group, Leaf_group, Drainage_group) %>% 
  unique() %>%
  right_join(t_results, by = c("Study_number", "SiteID")) -> t_results

```


```{r results='hide', message=FALSE, include=FALSE}
# swc and precipitation model
# results_orig <- pm_model_sum()
results <- results_orig
results %>% na.omit() -> results
results %>% rename(Study_number = sub_site.Study_number.1., SiteID = var_site.j., Measure_year = sub_site.Measure_Year.1. ) -> results
```


* Join and get the meta information
```{r}
# 152 GWC sites and 375 VWC sites
results %>% filter(var_swc == "GWC") %>% select(Study_number, SiteID) %>% unique
results %>% filter(var_swc == "VWC") %>% select(Study_number, SiteID) %>% unique

# use left join and get Ts_depth information
DGRsD_SWC %>% 
  select(Study_number, SiteID, SWC_Depth, Latitude, Longitude, MAP_Del, MAT_Del) %>% group_by(Study_number, SiteID) %>% 
  dplyr::summarise(SWCdepth = mean(SWC_Depth), Lat = mean(Latitude), Lon = mean(Longitude), MAP = mean(MAP_Del), MAT = mean(MAT_Del)) %>%
  right_join(results, by = c("Study_number", "SiteID")) -> results
 
# more numeric factors from SRDB_V4
SRDB %>% select(Study_number, SiteID, Elevation) %>% group_by (Study_number, SiteID) %>%
 dplyr:: summarise(Ele = mean(Elevation, na.rm = TRUE)) %>% 
  right_join(results, by = c("Study_number", "SiteID")) -> results

# more caetgorial factors from SRDB_V4
SRDB %>% select(Study_number, SiteID, Biome_group, Ecosystem_group, Leaf_group, Drainage_group) %>% 
  unique() %>%
  right_join(results, by = c("Study_number", "SiteID")) -> results

```

## Analyze whether air temperature is a good suggogate of soil temperature in soil respiration modeling
```{r}
# find out the ununsial relationship first
# for the SLR, find out those suties and sites with negative slope
t_results %>% filter (first_b < 0) %>% select (Study_number) %>% unique()
t_results %>% filter (poly_c > 0) %>% select (Study_number, poly_c) %>% unique() %>% arrange(desc(poly_c))

```



```{r}

# R2 difference if use SLR model
t_results %>% mutate(Dif_SLR = case_when(first_R2 - first_Tm_R2 < -0.1 ~ "Low"
                                          , first_R2 - first_Tm_R2 > 0.1 ~ "High"
                                          , TRUE ~ "Same")) %>% select(SiteID, first_R2, first_Tm_R2, Dif_SLR) %>% 
  count(Dif_SLR) %>% mutate (freq = round(n/sum(n) * 100,2)) %>% 
  ggplot(aes(x="", y=freq, fill = Dif_SLR)) +
  geom_bar(width = 1, stat = "identity") -> p1

p1 = p1 + coord_polar("y", start=0) + geom_text(aes(label = paste0(round(freq, 2), "%")), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values=c("violet", "orange", "gray")) +
  labs(x = NULL, y = NULL, fill = NULL, title = expression("(a) SLR"~R^{2}~"difference")) +
  theme_classic() + theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "bottom",
          plot.title = element_text(hjust = 0.5, color = "black"))


# plot the percentage of SLR p<=0.05 and p>0.05
t_results %>% mutate(p_group = ifelse(p_b < 0.05, "p<0.5", "p>=0.5")) %>% select(p_group) %>% 
  count(p_group) %>% mutate (freq = round(n/sum(n) * 100,2)) %>% 
  ggplot(aes(x="", y=freq, fill = p_group)) +
  geom_bar(width = 1, stat = "identity") -> p2

p2 = p2 + coord_polar("y", start=0) + geom_text(aes(label = paste0(round(freq, 2), "%")), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values=c("pink", "cyan")) +
  labs(x = NULL, y = NULL, fill = NULL, title = expression("(b) SLR Tsoil")) +
  theme_classic() + theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "bottom",
          plot.title = element_text(hjust = 0.5, color = "black"))


# plot the percentage of SLR p<=0.05 and p>0.05 using air temperature
t_results %>% mutate(p_group = ifelse(p_Tm_b < 0.05, "p<0.5", "p>=0.5")) %>% select(p_group) %>% 
  count(p_group) %>% mutate (freq = round(n/sum(n) * 100,2)) %>% 
  ggplot(aes(x="", y=freq, fill = p_group)) +
  geom_bar(width = 1, stat = "identity") -> p3

p3 = p3 + coord_polar("y", start=0) + geom_text(aes(label = paste0(round(freq, 2), "%")), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values=c("pink", "cyan")) +
  labs(x = NULL, y = NULL, fill = NULL, title = "(c) SLR Tair") +
  theme_classic() + theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "bottom",
          plot.title = element_text(hjust = 0.5, color = "black"))


# R2 difference if use Polynomial model
t_results %>% mutate(Dif_poly = case_when(poly_R2 - poly_Tm_R2 < -0.1 ~ "Low"
                                          , poly_R2 - poly_Tm_R2 > 0.1 ~ "High"
                                          , TRUE ~ "Same")) %>% select(SiteID, poly_R2, poly_Tm_R2, Dif_poly) %>% 
  count(Dif_poly) %>% mutate (freq = round(n/sum(n) * 100,2)) %>% 
  ggplot(aes(x="", y=freq, fill = Dif_poly)) +
  geom_bar(width = 1, stat = "identity") -> p4

p4 = p4 + coord_polar("y", start=0) + geom_text(aes(label = paste0(round(freq, 2), "%")), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values=c("violet", "orange", "gray")) +
  labs(x = NULL, y = NULL, fill = NULL, title = expression("(d) Poly"~R^{2}~"difference")) +
  theme_classic() + theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "bottom",
          plot.title = element_text(hjust = 0.5, color = "black"))

# plot the percentage of polynomial p<=0.05 and p>0.05
t_results %>% mutate(p_group = ifelse(p_poly_b <= 0.05 | p_c < 0.05, "p<0.5", "p>=0.5")) %>% select(p_group) %>% 
  count(p_group) %>% mutate (freq = round(n/sum(n) * 100,2)) %>% 
  ggplot(aes(x="", y=freq, fill = p_group)) +
  geom_bar(width = 1, stat = "identity") -> p5

p5 = p5 + coord_polar("y", start=0) + geom_text(aes(label = paste0(round(freq, 2), "%")), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values=c("pink", "cyan")) +
  labs(x = NULL, y = NULL, fill = NULL, title = "(e) Poly Tsoil") +
  theme_classic() + theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "bottom",
          plot.title = element_text(hjust = 0.5, color = "black"))


# plot the percentage of polynomial p<=0.05 and p>0.05
t_results %>% mutate(p_group = ifelse(p_poly_Tm_b <= 0.05 | p_Tm_c < 0.05, "p<0.5", "p>=0.5")) %>% select(p_group) %>% 
  count(p_group) %>% mutate (freq = round(n/sum(n) * 100,2)) %>% 
  ggplot(aes(x="", y=freq, fill = p_group)) +
  geom_bar(width = 1, stat = "identity") -> p6

p6 = p6 + coord_polar("y", start=0) + geom_text(aes(label = paste0(round(freq, 2), "%")), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values=c("pink", "cyan")) +
  labs(x = NULL, y = NULL, fill = NULL, title = "(f) Poly Tair") +
  theme_classic() + theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "bottom",
          plot.title = element_text(hjust = 0.5, color = "black"))

plot_grid(p1, p2, p3
          , p4, p5, p6, ncol = 3)

ggsave("outputs/Figure2.jpg", width = 8, height = 6, dpi = 300, units = "in")

# t_results %>% mutate(p_group = ifelse(p_poly_b <= 0.05 | p_c <= 0.05, "Yes", "No")) %>% select(SiteID, p_poly_b, p_c, p_group)
  
```



```{r}
# compare R2 of SLR and polynomial models
t_results %>% filter(p_b <= 0.05) %>% select (p_b, p_Tm_b, first_R2, first_Tm_R2) %>% nrow() -> obs1
t_results %>% filter(p_b <= 0.05) %>% select (p_b, p_Tm_b, first_R2, first_Tm_R2) %>% 
  ggplot(aes(x = paste0("Tsoil (n=", obs1, ")"), y = first_R2)) + geom_violin() + 
  geom_jitter(shape = 16, position = position_jitter(0.2), col = 'gray') +
  geom_boxplot(width = 0.1) + 
  ylab(expression(R^2~"(SLR)")) + 
  theme(axis.title.x = element_blank()
        , axis.text=element_text(size=12)
        , axis.title.y = element_text(size = 12)) -> p1

t_results %>% filter(p_Tm_b <= 0.05) %>% select (p_b, p_Tm_b, first_R2, first_Tm_R2) %>% nrow() -> obs2
t_results %>% filter(p_Tm_b <= 0.05) %>% select (p_b, p_Tm_b, first_R2, first_Tm_R2) %>% 
  ggplot(aes(x = paste0("Tair (n=", obs2, ")"), y = first_Tm_R2)) + geom_violin() + 
  geom_jitter(shape = 16, position = position_jitter(0.2), col = 'gray') +
  geom_boxplot(width = 0.1) + 
  ylab(expression(R^2~"(SLR)")) +
  theme(axis.title.x = element_blank()
        , axis.text = element_text(size = 12)
        , axis.title.y = element_text(size = 12)) -> p2


# compare R2 of Polynomial model
t_results %>% filter(p_c <= 0.05) %>% select (p_c, p_Tm_c, poly_R2, poly_Tm_R2) %>% nrow() -> obs3
t_results %>% filter(p_c <= 0.05) %>% select (p_c, p_Tm_c, poly_R2, poly_Tm_R2) %>% 
  ggplot(aes(x = paste0("Tsoil (n=", obs3, ")"), y = poly_R2)) + geom_violin() + 
  geom_jitter(shape = 16, position = position_jitter(0.2), col = 'gray') +
  geom_boxplot(width = 0.1) + 
  ylab(expression(R^2~"(Polynomial)")) + 
  theme(axis.title.x = element_blank()
        , axis.text=element_text(size=12)
        , axis.title.y = element_text(size = 12)) -> p3

t_results %>% filter(p_Tm_c <= 0.05) %>% select (p_c, p_Tm_c, poly_R2, poly_Tm_R2) %>% nrow() -> obs4
t_results %>% filter(p_Tm_c <= 0.05) %>% select (p_c, p_Tm_c, poly_R2, poly_Tm_R2) %>% 
  ggplot(aes(x = paste0("Tair (n=", obs4, ")"), y = poly_Tm_R2)) + geom_violin() + 
  geom_jitter(shape = 16, position = position_jitter(0.2), col = 'gray') +
  geom_boxplot(width = 0.1) + 
  ylab(expression(R^2~"(Polynomial)")) + 
  theme(axis.title.x = element_blank()
        , axis.text=element_text(size=12)
        , axis.title.y = element_text(size = 12)) -> p4

plot_grid(p1, p2, p3, p4, labels = c("(a)", "(b)", "(c)", "(d)")
          , vjust = 2, hjust = c(-4), ncol = 2 )

ggsave("outputs/Figure3-1.jpg", width = 8, height = 6, dpi = 300, units = "in")

t_results %>% filter(p_b <= 0.05) %>% select (p_b, p_Tm_b, first_R2, first_Tm_R2) %>% summarise(median(first_R2))
t_results %>% filter(p_Tm_b <= 0.05) %>% select (p_b, p_Tm_b, first_R2, first_Tm_R2) %>% summarise(median(first_Tm_R2))

t_results %>% filter(p_c <= 0.05) %>% select (p_c, p_Tm_c, poly_R2, poly_Tm_R2) %>% summarise(median(poly_Tm_R2))
t_results %>% filter(p_Tm_c <= 0.05) %>% select (p_c, p_Tm_c, poly_R2, poly_Tm_R2) %>% summarise(median(poly_Tm_R2))

```







* PCA to test the main factors
```{r}
# Use PCA to detect which environmental factor effect the surrogates of temperature
t_results %>% select(Lat, Lon, Ts_depth, MAP, MAT, Diff, Dif_group) %>% na.omit() -> sub_t_results
sub_t_results %>% dplyr::rename(Depth = Ts_depth) -> sub_t_results
prcomp(sub_t_results %>% select(Lat, Lon, Depth, MAP, MAT) %>% na.omit(),
             center = TRUE,
             scale. = TRUE) -> t_PCA

str(t_PCA)
summary(t_PCA)


g <- ggbiplot(t_PCA,
              obs.scale = 1,
              var.scale = 1,
              groups = sub_t_results$Dif_group,
              ellipse = T,
              circle = T,
              ellipse.prob = 0.9,
              varname.size =  5 )

g <- g + scale_color_discrete(name = '')

g <- g + theme_bw()

g <- g + theme(legend.direction = 'horizontal',
               legend.position = 'top')

print(g)
```




* Using box plot to show the comparison

```{r}
t_results %>% filter(Ecosystem_group == "Plantation") %>% na.omit() %>% nrow()
t_results %>% filter(Ecosystem_group == "Wetland") %>% na.omit() %>% nrow()

t_results %>% na.omit() %>% ggplot( aes(y = Diff, x = Biome_group) ) + geom_boxplot( width = 0.5) +
  theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(angle = 25, hjust = 1)) + 
  ylab(expression(SLR~R^{2}~difference)) -> p1

t_results %>% na.omit() %>% ggplot( aes(y = Diff, x = Ecosystem_group) ) + geom_boxplot( width = 0.5) +
  theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(angle = 25, hjust = 1)) +
  ylab(expression(SLR~R^{2}~difference)) -> p2

t_results %>% na.omit() %>% ggplot( aes(y = Diff, x = Leaf_group) ) + geom_boxplot( width = 0.5) +
  theme(axis.title.x = element_blank()) + ylab(expression(SLR~R^{2}~difference)) -> p3

t_results %>% na.omit() %>% ggplot( aes(y = Diff, x = Drainage_group) ) + geom_boxplot( width = 0.5) +
  theme(axis.title.x = element_blank()) + ylab(expression(SLR~R^{2}~difference)) -> p4

plot_grid(p1, p2, p3, p4, labels = c("(a)", "(b)", "(c)", "(d)")
          , ncol = 2
          , hjust = c(-3.25), vjust = c(2))

ggsave("outputs/Figure S2.jpg", dpi = 300, width = 8, height = 6, units = "in")

```



* Linear regression to test whether Latitude, longitude, elevation, MAT, and MAP effect the surrogates
```{r}
summary(lm(Diff~abs(Lat), data = sub_t_results))
summary(lm(Diff~Ts_depth, data = sub_t_results))
summary(lm(Diff~MAT, data = sub_t_results))
summary(lm(Diff~MAP, data = sub_t_results))
summary(lm(Diff~Ele, data = t_results))
summary(lm(Diff~Measure_year, data = t_results))

ggplot(data = sub_t_results, aes(Lon, Diff)) + geom_point(col = "gray", alpha=0.5) + geom_smooth(method = "lm") +
  labs(x = expression(Longitude), y = expression(SLR~R^{2}~difference))


ggplot(data = sub_t_results, aes(abs(Lat), Diff)) + geom_point(col = "gray", alpha=0.5) + geom_smooth(method = "lm") +
  labs(x = expression(Absolute~latitude), y = expression(SLR~R^{2}~difference)) -> p1

ggplot(data = sub_t_results, aes(Ts_depth, Diff)) + geom_point(col = "gray", alpha=0.5) + geom_smooth(method = "lm") +
  labs(x = expression(Tsoil~measure~depth~"(cm)"), y = expression(SLR~R^{2}~difference)) -> p2

ggplot(data = sub_t_results, aes(MAT, Diff)) + geom_point(col = "gray", alpha=0.5) + geom_smooth(method = "lm") +
  labs(x = expression(MAT~"("~degree~C~")"), y = expression(SLR~R^{2}~difference)) -> p3

ggplot(data = sub_t_results, aes(MAP, Diff)) + geom_point(col = "gray", alpha=0.5) + geom_smooth(method = "lm") +
  labs(x = expression(MAP~"(mm)"), y = expression(SLR~R^{2}~difference)) -> p4

ggplot(data = t_results, aes(Ele, Diff)) + geom_point(col = "gray", alpha=0.5) + geom_smooth(method = "lm") +
  labs(x = expression(Elevation~"(m)"), y = expression(SLR~R^{2}~difference)) -> p5

ggplot(data = t_results, aes(Measure_year, Diff)) + geom_point(col = "gray", alpha=0.5) + geom_smooth(method = "lm") +
  labs(x = expression(Measure~year), y = expression(SLR~R^{2}~difference)) -> p6


plot_grid (p1, p2, p3, p4,p5,p6, ncol = 2, labels = c("(a)", "(b)", "(c)", "(d)", "(e)", "(f)")
           , hjust = c(-3,-3,-3,-3,-3,-4), vjust = c(2))

ggsave("outputs/Figure6.jpg", dpi = 300, width = 8, height = 7, units = "in")
```



* Classfication modeling to study whether Biome, ecosystem_type, leaf_habit affect whether soil respiration has a significant relationship with Tsoil
```{r}
# classfication
fit <- rpart(Diff ~ Biome + Ecosystem_type + Leaf_habit + Soil_drainage,
             method="class", data=t_results %>% 
               select(Biome, Ecosystem_type, Leaf_habit, Soil_drainage, Diff) %>% na.omit())

printcp(fit) # display the results 
plotcp(fit) # visualize cross-validation results 
summary(fit) # detailed summary of splits

plot(fit, uniform=TRUE, 
   main="Classification Tree")

```


* Using logistic regression to study whether latitude, longitude, elevation, MAT, MAP effect whether soil respiration has significant relationship with Tsoil
```{r}
# logistic regression to test when p<0.05
t_results %>% mutate(p_group = ifelse(p_b <= 0.05, 1, 0)) %>% select(Study_number, SiteID, Lat, Lon, Ele, MAT, MAP, p_b, p_group, Biome) %>% 
  ggplot(aes(p_group, fill = Biome)) + geom_bar()

# logistic regression to test when p<0.05
t_results %>% mutate(p_group = ifelse(p_b <= 0.05, 1, 0)) %>% select(Study_number, SiteID, Lat, MAT, MAP, p_b, p_group, Biome) %>%
  na.omit() -> sub_t_results 
summary(glm(p_group ~ Lat  + MAT + MAP, data = sub_t_results, family = "binomial"))

```


* Calculate and show the threshold of Tsoil and Tair to soil respiration
* Note that only those sites with p<0.05 (significant) were included

```{r}
# calculate Tsoil threshold and Tair threshold
t_results %>% mutate (thd_Tsoil = -poly_b/(2*poly_c), thd_Tair = -poly_Tm_b/(2*poly_Tm_c)) -> t_results

```


* threhold of Tsoil and Tair
```{r}
# plot and comparison
t_results %>% filter(p_poly_b <= 0.05 & poly_c < 0 & thd_Tsoil < 65) %>% nrow() -> obs5
t_results %>% filter(p_poly_b <= 0.05 & poly_c < 0 & thd_Tsoil < 65) %>% 
  ggplot(aes(x = paste0("Tsoil (n=", obs5, ")"), y = thd_Tsoil)) + geom_violin() + 
  geom_jitter(shape = 16, position = position_jitter(0.2), col = 'gray') +
  geom_boxplot(width = 0.1) + 
  ylab(expression(Threshold~"(Tsoil"~degree~C~")")) + 
  theme(axis.title.x = element_blank()
        , axis.text=element_text(size=10)
        , axis.title.y = element_text(size = 10)) -> p1

# threshold of Tair

t_results %>% filter(p_poly_Tm_b <= 0.05 & poly_Tm_c < 0 & thd_Tair < 65) %>% nrow() -> obs7
t_results %>% filter(p_poly_Tm_b <= 0.05 & poly_Tm_c < 0 & thd_Tair < 65) %>% 
  ggplot(aes(x = paste0("Tair (n=", obs7, ")"), y = thd_Tair)) + geom_violin() + 
  geom_jitter(shape = 16, position = position_jitter(0.2), col = 'gray') +
  geom_boxplot(width = 0.1) + 
  ylab(expression(Threshold~"(Tair"~degree~C~")")) + 
  theme(axis.title.x = element_blank()
        , axis.text=element_text(size=10)
        , axis.title.y = element_text(size = 10)) -> p2

plot_grid(p1, p2, labels = c( "(a)", "(b)" )
          , vjust = 2, hjust = c(-3), ncol = 1 )

ggsave("outputs/Figure3-2.jpg", width = 4, height = 4, dpi = 300, units = "in")

t_results %>% filter(p_poly_b <= 0.05 & poly_c < 0 & thd_Tsoil < 65) %>% summarise(median(thd_Tsoil))
t_results %>% filter(p_poly_Tm_b <= 0.05 & poly_Tm_c < 0 & thd_Tair < 65) %>% summarise(median(thd_Tair))

```









## Analyze whether soil water content is a good suggogate of soil moisture in soil respiration modeling

## Analyze whether soil water content is a good suggogate of soil moisture in soil respiration modeling


```{r}

# Rs vs GWC, if put all data together, show no relationship
# due to to much site to site variability
ggplot(subset(DGRsD, SWC_Type == "GWC")) +
  aes(x = SWC, y = RS_Norm) +
  geom_hex ( ) +
  scale_fill_gradient(low = "gray", high = "blue") +
  # geom_smooth(method = "lm") + 
  xlab(expression(SWC~"("~"%"~")")) +
  ylab(expression(R[s]~g~C~m^{-2}))

# DGRsD %>% filter(SWC_Type == "VWC") %>% lm(RS_Norm ~ SWC) %>%  summary ()
```



```{r}
# Rs vs VWC by climate region, show very weak relationship
# linear repression and statistics
SLR_sum <- function (sdata, climate) {
  sdata %>% dplyr::filter(SWC_Type == "VWC" & !is.na(DGRsD$SWC) & DGRsD$Climate_type == climate) -> subdata 
  lm(RS_Norm ~ SWC, data = subdata) %>% summary() %>% print()
}

SLR_sum(DGRsD, "(A) Tropic")
SLR_sum(DGRsD, "(B) Arid")
SLR_sum(DGRsD, "(C) Temperate")
SLR_sum(DGRsD, "(D) Continental")

# library(dplyr)

DGRsD %>% dplyr::filter(SWC_Type == "VWC" & !is.na(DGRsD$SWC)) %>%  
  ggplot() +
  aes(x = SWC, y = RS_Norm) +
  geom_hex ( ) +
  scale_fill_gradient(low = "gray", high = "blue") +
  # geom_smooth(method = "lm") + 
  xlab(expression(SWC~"("~"%"~")")) +
  ylab(expression(R[s]~"("~g~C~m^{-2}~day^{-1}~")")) +
  facet_grid(rows = vars(Climate_type)) + 
  geom_smooth(method = "lm") +
  # annotate("text", label = c(expression(R^2~"=0.00 (p>0.05)"), expression(R^2~"=0.11 (p<0.01)"),
  #                            expression(R^2~"=0.00 (p>0.05)"), expression(R^2~"=0.05 (p>0.05)")  )
  #          , size = 4, x = 60, y = 20) +
  theme(legend.position = "none") -> plot_vwc


dat_text <- data.frame(
  label = c("R2 = 0.01 (p<0.05)", "R2 = 0.11 (p<0.01)", "R2 = 0.00 (p>0.05)", "R2 = 0.05 (p>0.05)"),
  Climate_type   = c("(A) Tropic", "(B) Arid", "(C) Temperate", "(D) Continental") )

plot_vwc <- plot_vwc + geom_text(
  data    = dat_text,
  mapping = aes(x = -Inf, y = -Inf, label = label),
  hjust   = -3,
  vjust   = -5
)

# Rs vs Pm_Del 
SLR_Pm <- function (sdata, climate) {
  sdata %>% filter(SWC_Type == "VWC" & !is.na(DGRsD$Pm_Del) & DGRsD$Climate_type == climate) -> subdata 
  lm(RS_Norm ~ Pm_Del, data = subdata) %>% summary() %>% print()
}

SLR_Pm(DGRsD, "(A) Tropic")
SLR_Pm(DGRsD, "(B) Arid")
SLR_Pm(DGRsD, "(C) Temperate")
SLR_Pm(DGRsD, "(D) Continental")


DGRsD %>% filter(SWC_Type == "VWC" & !is.na(DGRsD$SWC) & !is.na(DGRsD$Pm_Del) ) %>%
  ggplot() +
  aes(x = Pm_Del, y = RS_Norm) +
  geom_hex ( ) +
  scale_fill_gradient(low = "gray", high = "blue") +
  # geom_smooth(method = "lm") + 
  xlab(expression(Preiptation~"(mm)")) +
  ylab(expression(R[s]~"("~g~C~m^{-2}~day^{-1}~")")) +
  facet_grid(rows = vars(Climate_type)) +
  geom_smooth(method = "lm") +
  # annotate("text", label = c(expression(R^2~"=0.01 (p<0.05)"), expression(R^2~"=0.02 (p<0.01)"),
  #                            expression(R^2~"=0.05 (p<0.05)"), expression(R^2~"=0.12 (p<0.01)")  )
  #          , size = 4, x = 750, y = 20) +
  theme(legend.position = "none") -> plot_pm

dat_text <- data.frame(
  label = c("R2 = 0.01 (p<0.05)", "R2 = 0.02 (p<0.01)", "R2 = 0.05 (p<0.05)", "R2 = 0.12 (p<0.01)"),
  Climate_type   = c("(A) Tropic", "(B) Arid", "(C) Temperate", "(D) Continental") )

plot_pm <- plot_pm + geom_text(
  data    = dat_text,
  mapping = aes(x = -Inf, y = -Inf, label = label),
  hjust   = -3,
  vjust   = -5
)


# tag_facet(plot_pm)
# ?tag_facet()

plot_grid(plot_vwc, plot_pm) 
# grid.arrange(plot_vwc, plot_pm, ncol = 2)

# plot_vwc
ggsave("outputs/Figure S4.jpg", width = 8, height = 7, dpi = 300, units = "in")

# plot_pm
# ggsave("outputs/Figure S4-2.jpg", width = 4, height = 7, dpi = 300, units = "in")
```


* Using box plot to show the comparison

```{r}


results %>% na.omit() %>% ggplot( aes(y = poly_R2 - poly_pm_R2, x = Biome_group) ) + geom_boxplot( width = 0.5) +
  theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(angle = 25, hjust = 1)) + 
  ylab(expression(Poly~R^{2}~difference)) -> p1

results %>% na.omit() %>% ggplot( aes(y = poly_R2 - poly_pm_R2, x = Ecosystem_group) ) + geom_boxplot( width = 0.5) +
  theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(angle = 25, hjust = 1)) +
  ylab(expression(Poly~R^{2}~difference)) -> p2

results %>% na.omit() %>% ggplot( aes(y = poly_R2 - poly_pm_R2, x = Leaf_group) ) + geom_boxplot( width = 0.5) +
  theme(axis.title.x = element_blank()) + ylab(expression(Poly~R^{2}~difference)) -> p3

results %>% na.omit() %>% ggplot( aes(y = poly_R2 - poly_pm_R2, x = Drainage_group) ) + geom_boxplot( width = 0.5) +
  theme(axis.title.x = element_blank()) + ylab(expression(Poly~R^{2}~difference)) -> p4

plot_grid(p1, p2, p3, p4, labels = c("(a)", "(b)", "(c)", "(d)")
          , ncol = 2
          , hjust = c(-3.25), vjust = c(2))

ggsave("outputs/Figure S5.jpg", dpi = 300, width = 8, height = 6, units = "in")

```


* Linear regression to test whether Latitude, longitude, elevation, MAT, and MAP effect the surrogates
```{r}

results %>% select(Lat, Lon, SWCdepth, MAP, MAT, poly_R2, poly_pm_R2) %>% na.omit() %>% mutate(Diff = poly_R2 - poly_pm_R2) -> sub_results

summary(lm(Diff~abs(Lat), data = sub_results))
summary(lm(Diff~SWCdepth, data = sub_results))
summary(lm(Diff~MAT, data = sub_results))
summary(lm(Diff~MAP, data = sub_results))
results %>%  mutate(Diff = poly_R2 - poly_pm_R2) -> results
summary(lm(Diff~Ele, data = results))
summary(lm(Diff~Measure_year, data = results))

ggplot(data = sub_results, aes(Lon, Diff)) + geom_hex() + 
  scale_fill_gradient(low = "gray", high = "blue") +
  geom_smooth(method = "lm") +
  labs(x = expression(Longitude), y = expression(Poly~R^{2}~difference))


ggplot(data = sub_results, aes(abs(Lat), Diff)) + geom_point(col="gray", alpha=0.5) + 
  # scale_fill_gradient(low = "gray", high = "blue") + theme(legend.position = "none") +
  geom_smooth(method = "lm") +
  labs(x = expression(Absolute~latitude), y = expression(Poly~R^{2}~difference)) -> p1

ggplot(data = sub_results, aes(SWCdepth, Diff)) + geom_point(col="gray", alpha=0.5) + 
  # scale_fill_gradient(low = "gray", high = "blue") + theme(legend.position = "none") +
  geom_smooth(method = "lm") +
  labs(x = expression(SWC~measure~depth~"(cm)"), y = expression(Poly~R^{2}~difference)) -> p2

ggplot(data = sub_results, aes(MAT, Diff)) + geom_point(col="gray", alpha=0.5) + 
  # scale_fill_gradient(low = "gray", high = "blue") + theme(legend.position = "none") +
  geom_smooth(method = "lm") +
  labs(x = expression(MAT~"("~degree~C~")"), y = expression(Poly~R^{2}~difference)) -> p3

ggplot(data = sub_results %>% filter(MAP < 5000), aes(MAP, Diff)) + geom_point(col="gray", alpha=0.5) + 
  # scale_fill_gradient(low = "gray", high = "blue") + theme(legend.position = "none") +
  geom_smooth(method = "lm") +
  labs(x = expression(MAP~"(mm)"), y = expression(Poly~R^{2}~difference)) -> p4

results %>% filter(Ele <= 5000) %>% mutate(Diff = poly_R2 - poly_pm_R2) %>%  
  ggplot(aes(Ele, Diff)) + geom_point(col="gray", alpha=0.5) +
  # scale_fill_gradient(low = "gray", high = "blue") + theme(legend.position = "none") +
  geom_smooth(method = "lm") +
  labs(x = expression(Elevation~"(m)"), y = expression(Poly~R^{2}~difference)) -> p5

ggplot(data = results %>% mutate(Diff = poly_R2 - poly_pm_R2), aes(Measure_year, Diff)) + geom_point(col="gray", alpha=0.5) + 
  # scale_fill_gradient(low = "gray", high = "blue") + theme(legend.position = "none") +
  geom_smooth(method = "lm") +
  labs(x = expression(Measure~year), y = expression(Poly~R^{2}~difference)) -> p6

plot_grid (p1, p2, p3, p4,p5, p6, ncol = 2, labels = c("(a)", "(b)", "(c)", "(d)", "(e)", "(f)")
           , hjust = c(-3,-3,-3,-3,-3,-4), vjust = c(2))

ggsave("outputs/Figure7.jpg", dpi = 300, width = 8, height = 7, units = "in")
```



```{r}
results$first_R2 %>% mean()
results$poly_R2 %>% mean()
results$first_pm_R2 %>% mean()
results$poly_pm_R2 %>% mean(na.rm=T)
```



* Test and plot the percentage of p<0.05
```{r}
# Precipitation as a surrogate of SWC
# R2 difference if use SLR model

results %>% mutate(Dif_SLR = case_when(first_R2 - first_pm_R2 < -0.1 ~ "Low"
                                          , first_R2 - first_pm_R2 > 0.1 ~ "High"
                                          , TRUE ~ "Same")) %>% select(SiteID, first_R2, first_pm_R2, Dif_SLR) %>% 
  count(Dif_SLR) %>% mutate (freq = round(n/sum(n) * 100,2)) %>% 
  ggplot(aes(x="", y=freq, fill = Dif_SLR)) +
  geom_bar(width = 1, stat = "identity") -> p1

p1 = p1 + coord_polar("y", start=0) + geom_text(aes(label = paste0(round(freq, 2), "%")), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values=c("violet", "orange", "gray")) +
  labs(x = NULL, y = NULL, fill = NULL, title = expression("(a) SLR"~R^{2}~"difference")) +
  theme_classic() + theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "bottom",
          plot.title = element_text(hjust = 0.5, color = "black"))


# plot the percentage of SLR p<=0.05 and p>0.05
results %>% mutate(p_group = ifelse(p_b < 0.05, "p<0.5", "p>=0.5")) %>% select(p_group) %>% 
  count(p_group) %>% mutate (freq = round(n/sum(n) * 100,2)) %>% 
  ggplot(aes(x="", y=freq, fill = p_group)) +
  geom_bar(width = 1, stat = "identity") -> p2

p2 = p2 + coord_polar("y", start=0) + geom_text(aes(label = paste0(round(freq, 2), "%")), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values=c("pink", "cyan")) +
  labs(x = NULL, y = NULL, fill = NULL, title = expression("(b) SLR SWC")) +
  theme_classic() + theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "bottom",
          plot.title = element_text(hjust = 0.5, color = "black"))


# plot the percentage of SLR p<=0.05 and p>0.05 using precipitation
results %>% mutate(p_group = ifelse(p_pm_b < 0.05, "p<0.5", "p>=0.5")) %>% select(p_group) %>% 
  count(p_group) %>% mutate (freq = round(n/sum(n) * 100,2)) %>% 
  ggplot(aes(x="", y=freq, fill = p_group)) +
  geom_bar(width = 1, stat = "identity") -> p3

p3 = p3 + coord_polar("y", start=0) + geom_text(aes(label = paste0(round(freq, 2), "%")), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values=c("pink", "cyan")) +
  labs(x = NULL, y = NULL, fill = NULL, title = "(c) SLR precipitation") +
  theme_classic() + theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "bottom",
          plot.title = element_text(hjust = 0.5, color = "black"))


# R2 difference if use Polynomial model
results %>% mutate(Dif_poly = case_when(poly_R2 - poly_pm_R2 < -0.1 ~ "Low"
                                          , poly_R2 - poly_Tm_R2 > 0.1 ~ "High"
                                          , TRUE ~ "Same")) %>% select(SiteID, poly_R2, poly_pm_R2, Dif_poly) %>% 
  count(Dif_poly) %>% mutate (freq = round(n/sum(n) * 100,2)) %>% 
  ggplot(aes(x="", y=freq, fill = Dif_poly)) +
  geom_bar(width = 1, stat = "identity") -> p4

p4 = p4 + coord_polar("y", start=0) + geom_text(aes(label = paste0(round(freq, 2), "%")), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values=c("violet", "orange", "gray")) +
  labs(x = NULL, y = NULL, fill = NULL, title = expression("(d) Poly"~R^{2}~"difference")) +
  theme_classic() + theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "bottom",
          plot.title = element_text(hjust = 0.5, color = "black"))

# plot the percentage of polynomial p<=0.05 and p>0.05
results %>% mutate(p_group = ifelse(p_poly_b <= 0.05 | p_c < 0.05, "p<0.5", "p>=0.5")) %>% select(p_group) %>% 
  count(p_group) %>% mutate (freq = round(n/sum(n) * 100,2)) %>% 
  ggplot(aes(x="", y=freq, fill = p_group)) +
  geom_bar(width = 1, stat = "identity") -> p5

p5 = p5 + coord_polar("y", start=0) + geom_text(aes(label = paste0(round(freq, 2), "%")), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values=c("pink", "cyan")) +
  labs(x = NULL, y = NULL, fill = NULL, title = "(e) Poly SWC") +
  theme_classic() + theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "bottom",
          plot.title = element_text(hjust = 0.5, color = "black"))


# plot the percentage of polynomial p<=0.05 and p>0.05
results %>% mutate(p_group = ifelse(p_poly_Tm_b <= 0.05 | p_pm_c < 0.05, "p<0.5", "p>=0.5")) %>% select(p_group) %>% 
  count(p_group) %>% mutate (freq = round(n/sum(n) * 100,2)) %>% 
  ggplot(aes(x="", y=freq, fill = p_group)) +
  geom_bar(width = 1, stat = "identity") -> p6

p6 = p6 + coord_polar("y", start=0) + geom_text(aes(label = paste0(round(freq, 2), "%")), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values=c("pink", "cyan")) +
  labs(x = NULL, y = NULL, fill = NULL, title = "(f) Poly precipitation") +
  theme_classic() + theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "bottom",
          plot.title = element_text(hjust = 0.5, color = "black"))

plot_grid(p1, p2, p3
          , p4, p5, p6, ncol = 3)

ggsave("outputs/Figure4.jpg", width = 8, height = 6, dpi = 300, units = "in")

results %>% mutate(p_group = ifelse(p_poly_b <= 0.05 | p_c <= 0.05, "Yes", "No")) %>% select(SiteID, p_poly_b, p_c, p_group)
  
```




```{r}
model_p_plot()
```



```{r}
# compare R2 of SLR 
results %>% filter(p_b <= 0.05) %>% select (p_b, p_pm_b, first_R2, first_pm_R2) %>% nrow() -> obs1
results %>% filter(p_b <= 0.05) %>% select (p_b, p_pm_b, first_R2, first_pm_R2) %>% 
  ggplot(aes(x = paste0("SWC (n=", obs1, ")"), y = first_R2)) + geom_violin() + 
  geom_jitter(shape = 16, position = position_jitter(0.2), col = 'gray') +
  geom_boxplot(width = 0.1) + 
  ylab(expression(R^2~"(SLR)")) + 
  theme(axis.title.x = element_blank()
        , axis.text=element_text(size=12)
        , axis.title.y = element_text(size = 12)) -> p1

results %>% filter(p_pm_b <= 0.05) %>% select (p_b, p_pm_b, first_R2, first_pm_R2) %>% nrow() -> obs2
results %>% filter(p_pm_b <= 0.05) %>% select (p_b, p_pm_b, first_R2, first_pm_R2) %>% 
  ggplot(aes(x = paste0("Precipitation (n=", obs2, ")"), y = first_pm_R2)) + geom_violin() + 
  geom_jitter(shape = 16, position = position_jitter(0.2), col = 'gray') +
  geom_boxplot(width = 0.1) + 
  ylab(expression(R^2~"(SLR)")) +
  theme(axis.title.x = element_blank()
        , axis.text = element_text(size = 12)
        , axis.title.y = element_text(size = 12)) -> p2


# compare R2 of Polynomial model
results %>% filter(p_c <= 0.05) %>% select (p_c, p_pm_c, poly_R2, poly_pm_R2) %>% nrow() -> obs3
results %>% filter(p_c <= 0.05) %>% select (p_c, p_pm_c, poly_R2, poly_pm_R2) %>% 
  ggplot(aes(x = paste0("SWC (n=", obs3, ")"), y = poly_R2)) + geom_violin() + 
  geom_jitter(shape = 16, position = position_jitter(0.2), col = 'gray') +
  geom_boxplot(width = 0.1) + 
  ylab(expression(R^2~"(Polynomial)")) + 
  theme(axis.title.x = element_blank()
        , axis.text=element_text(size=12)
        , axis.title.y = element_text(size = 12)) -> p3

results %>% filter(p_pm_c <= 0.05) %>% select (p_c, p_pm_c, poly_R2, poly_pm_R2) %>% nrow() -> obs4
results %>% filter(p_pm_c <= 0.05) %>% select (p_c, p_pm_c, poly_R2, poly_pm_R2) %>% 
  ggplot(aes(x = paste0("Precipitation (n=", obs4, ")"), y = poly_pm_R2)) + geom_violin() + 
  geom_jitter(shape = 16, position = position_jitter(0.2), col = 'gray') +
  geom_boxplot(width = 0.1) + 
  ylab(expression(R^2~"(Polynomial)")) + 
  theme(axis.title.x = element_blank()
        , axis.text=element_text(size=12)
        , axis.title.y = element_text(size = 12)) -> p4

plot_grid(p1, p2, p3, p4, labels = c("(a)", "(b)", "(c)", "(d)")
          , vjust = 2, hjust = c(-4), ncol = 2 )

ggsave("outputs/Figure5-1.jpg", width = 8, height = 6, dpi = 300, units = "in")

results %>% filter(p_b <= 0.05) %>% select (p_b, p_pm_b, first_R2, first_pm_R2) %>% summarise(median(first_R2))
results %>% filter(p_pm_b <= 0.05) %>% select (p_b, p_pm_b, first_R2, first_pm_R2) %>% summarise(median(first_pm_R2))

results %>% filter(p_c <= 0.05) %>% select (p_c, p_pm_c, poly_R2, poly_pm_R2) %>% summarise(median(poly_R2))
results %>% filter(p_pm_c <= 0.05) %>% select (p_c, p_pm_c, poly_R2, poly_pm_R2) %>% summarise(median(poly_pm_R2))
```



```{r}
# Calculate threshold 
results %>% mutate(thd_swc = -poly_b/(2*poly_c), thd_pm = -poly_pm_b/(2*poly_pm_c)) -> results
```


```{r}
# threshold of GWC
results %>% filter(p_poly_b <= 0.05 & p_c <= 0.05 & var_swc == "GWC") %>% nrow() -> obs5
results %>% filter(p_poly_b <= 0.05 & p_c <= 0.05 & var_swc == "GWC") %>% 
  ggplot(aes(x = paste0("GWC (n=", obs5, ")"), y = thd_swc)) + geom_violin() + 
  geom_jitter(shape = 16, position = position_jitter(0.2), col = 'gray') +
  geom_boxplot(width = 0.1) + 
  ylab(expression(Threshold~"(GWC %)")) + 
  theme(axis.title.x = element_blank()
        , axis.text=element_text(size=10)
        , axis.title.y = element_text(size = 10)) -> p1

# threshold of VWC
results %>% filter(p_poly_b <= 0.05 & p_c <= 0.05 & var_swc == "VWC") %>% nrow() -> obs6
results %>% filter(p_poly_b <= 0.05 & p_c <= 0.05 & var_swc == "VWC") %>% 
  ggplot(aes(x = paste0("VWC (n=", obs6, ")"), y = thd_swc)) + geom_violin() + 
  geom_jitter(shape = 16, position = position_jitter(0.2), col = 'gray') +
  geom_boxplot(width = 0.1) + 
  ylab(expression(Threshold~"(VWC %)")) + 
  theme(axis.title.x = element_blank()
        , axis.text=element_text(size=10)
        , axis.title.y = element_text(size = 10)) -> p2


# threshold of Pm

results %>% filter(p_poly_pm_b <= 0.05 & p_pm_c <= 0.05) %>% nrow() -> obs7
results %>% filter(p_poly_pm_b <= 0.05 & p_pm_c <= 0.05) %>% 
  ggplot(aes(x = paste0("Precipitation (n=", obs7, ")"), y = thd_pm)) + geom_violin() + 
  geom_jitter(shape = 16, position = position_jitter(0.2), col = 'gray') +
  geom_boxplot(width = 0.1) + 
  ylab(expression(Threshold~"(Pm, mm"~month^{-1}~")")) + 
  theme(axis.title.x = element_blank()
        , axis.text=element_text(size=10)
        , axis.title.y = element_text(size = 10)) -> p3

plot_grid(p1, p2, p3, labels = c( "(a)", "(b)", "(c)" )
          , vjust = 2, hjust = c(-4), ncol = 1 )

ggsave("outputs/Figure5-2.jpg", width = 4, height = 6, dpi = 300, units = "in")

```

```{r}
# threshold of VWC to relative water content = theta / poroacity
results %>% filter(p_poly_b <= 0.05 & p_c <= 0.05 & var_swc == "GWC") %>% select(thd_swc) %>% summarise(median(thd_swc) ) * 1.2313
results %>% filter(p_poly_b <= 0.05 & p_c <= 0.05 & var_swc == "VWC") %>% select(thd_swc) %>% summarise(median(thd_swc) )
results %>% filter(p_poly_b <= 0.05 & p_c <= 0.05 & var_swc == "VWC") %>% select(thd_swc) %>% summarise(median(thd_swc) / (1.2313/2.65))
results %>% filter(p_poly_pm_b <= 0.05 & p_pm_c <= 0.05) %>% select(thd_pm) %>% summarise(mpm = median(thd_pm)) * 12
DGRsD_SWC %>% filter (SWC_Type == "VWC") %>% nrow()
DGRsD_SWC %>% filter (SWC_Type == "GWC") %>% nrow()

```


```{r}
# Rs vs Pm_Del 
SLR_T <- function (sdata, climate) {
  sdata %>% filter(!is.na(sdata$Tsoil) & !is.na(sdata$Tm_Del) & sdata$Climate_type == climate) -> subdata 
  lm(Tsoil ~ Tm_Del, data = subdata) %>% summary() %>% print()
}

SLR_T(DGRsD, "(A) Tropic")
SLR_T(DGRsD, "(B) Arid")
SLR_T(DGRsD, "(C) Temperate")
SLR_T(DGRsD, "(D) Continental")

DGRsD %>% select(Tsoil, Tm_Del, Climate_type) %>% na.omit() %>% 
  ggplot(aes(Tm_Del, Tsoil)) + geom_point(col = "gray", alpha=0.25) + 
  geom_smooth(method = "lm") +
  facet_grid(row = vars(Climate_type)) +
  labs (x = expression (Air~temperature~"("~degree~C~")")
        , y = expression (Soil~temperature~"("~degree~C~")") ) -> p1

SLR_SWCP <- function (sdata, climate) {
  sdata %>% filter(!is.na(sdata$SWC) & !is.na(sdata$Pm_Del) & sdata$Climate_type == climate) -> subdata 
  lm(SWC ~ Pm_Del, data = subdata) %>% summary() %>% print()
}

SLR_SWCP(DGRsD, "(A) Tropic")
SLR_SWCP(DGRsD, "(B) Arid")
SLR_SWCP(DGRsD, "(C) Temperate")
SLR_SWCP(DGRsD, "(D) Continental")

DGRsD %>% select(SWC, Pm_Del, Climate_type) %>% na.omit() %>% 
  ggplot(aes(Pm_Del, SWC)) + geom_point(col = "gray", alpha = 0.25) + 
  geom_smooth(method = "lm") +
  facet_grid(row = vars(Climate_type)) +
  labs (x = expression (Precipitation~"("~mm~month^{-1}~")")
        , y = expression (SWC~"(%)") ) -> p2

plot_grid(p1, p2, ncol = 2
          , hjust = c(-2), vjust = c(2))

ggsave("outputs/figure s3.jpg", width = 8, height = 7, dpi = 300, units = "in")

```

```{r}
# find out the sites both Pm and SWC can explian RS variability
results %>% filter (p_b < 0.05) %>% nrow
results %>% filter (p_pm_b < 0.05) %>% nrow
results %>% filter (p_b < 0.05 & p_pm_b < 0.05) %>% select(first_R2) %>% summarise(median(first_R2))
results %>% filter (p_b < 0.05 & p_pm_b < 0.05) %>% select(first_pm_R2) %>% summarise(median(first_pm_R2))
results %>% select(Study_number) %>% unique() %>% nrow()

# plot only those 
results %>% filter(p_b <= 0.05) %>% select (p_b, p_pm_b, first_R2, first_pm_R2) %>% nrow() -> obs1
results %>% filter(p_b <= 0.05) %>% select (p_b, p_pm_b, first_R2, first_pm_R2) %>% 
  ggplot(aes(x = paste0("SWC (n=", obs1, ")"), y = first_R2)) + geom_violin() + 
  geom_jitter(shape = 16, position = position_jitter(0.2), col = 'gray') +
  geom_boxplot(width = 0.1) + 
  ylab(expression(R^2~"(SLR)")) + 
  theme(axis.title.x = element_blank()
        , axis.text=element_text(size=12)
        , axis.title.y = element_text(size = 12))

results %>% filter(p_pm_b <= 0.05) %>% select (p_b, p_pm_b, first_R2, first_pm_R2) %>% nrow() -> obs2
results %>% filter(p_pm_b <= 0.05) %>% select (p_b, p_pm_b, first_R2, first_pm_R2) %>% 
  ggplot(aes(x = paste0("Precipitation (n=", obs2, ")"), y = first_pm_R2)) + geom_violin() + 
  geom_jitter(shape = 16, position = position_jitter(0.2), col = 'gray') +
  geom_boxplot(width = 0.1) + 
  ylab(expression(R^2~"(SLR)")) +
  theme(axis.title.x = element_blank()
        , axis.text = element_text(size = 12)
        , axis.title.y = element_text(size = 12))

```

```{r}
# output modeling results as csv
# write.csv(t_results, "outputs/temp_surrogate.csv", row.names = FALSE)
# write.csv(results, "outputs/swc_surrogate.csv", row.names = FALSE)
```



```{r Figure S6}
# See how SWC measure_years
DGRsD_SWC %>% 
  filter(stringr::str_detect(SWC_Comments, 'Oven') )

DGRsD_SWC %>% mutate( SWC_Method = case_when(stringr::str_detect(SWC_Comments, 'Oven|dried|core') ~ "OvenDry" 
                                             , SWC_Comments==""|SWC_Comments=="%"|stringr::str_detect(SWC_Comments, 'Helvey') ~ "Not reported"
                                             , TRUE ~ "TDR") ) %>% select(SWC, SWC_Comments, SWC_Method, Measure_Year) %>% 
  filter(Measure_Year < 1980 & SWC_Method=="TDR")

DGRsD_SWC %>% mutate( SWC_Method = case_when(stringr::str_detect(SWC_Comments, 'Oven|dried|core') ~ "OvenDry" 
                                             , SWC_Comments==""|SWC_Comments=="%"|stringr::str_detect(SWC_Comments, 'Helvey') ~ "Not reported"
                                             , TRUE ~ "TDR") ) %>% count(SWC_Method)

DGRsD_SWC %>% mutate( SWC_Method = case_when(stringr::str_detect(SWC_Comments, 'Oven|dried') ~ "OvenDry" 
                                             , SWC_Comments==""|stringr::str_detect(SWC_Comments, 'Helvey') ~ "Not reported"
                                             , TRUE ~ "TDR") ) %>% filter(SWC_Method=="TDR") %>% count(SWC_Comments)

# Figure S6
DGRsD_SWC %>% mutate( SWC_Method = case_when(stringr::str_detect(SWC_Comments, 'Oven|dried|core') ~ "Oven dry" 
                                             , SWC_Comments==""|SWC_Comments=="%"|stringr::str_detect(SWC_Comments, 'Helvey') ~ "Not reported"
                                             , TRUE ~ "TDR") ) %>% 
  ggplot(aes(Measure_Year, SWC)) + 
  # geom_point(col="gray", alpha = 0.25) + 
  geom_hex ( ) +
  scale_fill_gradient(low = "gray", high = "blue") +
  facet_grid(row = vars(SWC_Method)) +
  xlab(expression(Measure~year)) +
  ylab(expression(SWC~"(%)"))

ggsave("outputs/figure S6.jpg", width = 6, height = 6, units = "in", dpi = 300)

# Figure S7
DGRsD_SWC %>% mutate( SWC_Method = case_when(stringr::str_detect(SWC_Comments, 'Oven|dried|core') ~ "Oven dry" 
                                             , SWC_Comments==""|SWC_Comments=="%"|stringr::str_detect(SWC_Comments, 'Helvey') ~ "Not reported"
                                             , TRUE ~ "TDR") ) %>% select(Study_number, SiteID, SWC_Method) %>% unique() %>% 
  right_join(results, by = c("Study_number", "SiteID")) -> DGRsD_SWC_MMethod

DGRsD_SWC_MMethod %>% 
  ggplot(aes(Ele, Diff)) + 
  # geom_point(col="gray", alpha = 0.25) + 
  geom_hex ( ) +
  scale_fill_gradient(low = "gray", high = "blue") +
  facet_grid(row = vars(SWC_Method)) +
  xlab(expression(Elevation~"(m)")) +
  ylab(expression(R^{2}~"difference"))

```


